True versus apparent shapes of astrophysical bow shocks

### 2017-05-26 papers/Paper1/sec-quadrics-observations.tex
Discussion and tweaking of box plot figure

We developed a methodology for automatically tracing the arcs as follows:
\begin{enumerate}[1.]
\item Calculate arrays of celestial coordinates, \(C\), for each pixel
  of the image.
\item Using the central source coordinates, \(C_0\) and nominal
  bowshock radius, \(R_0\) from \citet{Kobulnicky:2016a}, construct a
  pixel mask that includes only those pixels with separations from the
  source that satisfy \(\frac12 R_0 \le |C - C_0| \le 3 R_0\).  This mask
  will be used for all subsequent operations, which serves to help
  avoid confusion from the star itself and other bright sources in the
  field of view.
\item Define a ``step-back'' point, \(C_1\), which is located at a
  separation \(2 R_0\) from the source, but in the opposite direction
  from the apex of the bow shock. That is, along a position angle
  180\degr{} from the nominal position angle, \(\text{PA}_0\), of the
  bow shock axis.  This point is at one end of the orange line shown
  superimposed on the bow shock images in
  Figure~\ref{fig:mipsgal-examples}.
\item Looping over a grid of 50 position angles, \(\text{PA}_k\),
  within \(\pm 60\degr\) of \(\text{PA}_0\), estimate the location of
  the arc along rays cast from the step-back point, using two
  different methods:
  \begin{enumerate}[(a)]
  \item The pixel with the peak brightness, with coordinates
    \(C_{k,\text{peak}}\) (red dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \item The mean brightness-weighted separation from \(C_1\), with
    coordinates \(C_{k,\text{mean}}\) (light blue dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \end{enumerate}
  For each \(\text{PA}_k\) in the grid, the calculation is performed
  over only those pixels that satisfy
  \(|\text{PA}(C, C_1) - \text{PA}_k| < \frac12 \delta\theta\), where
  \(\delta\theta = 120/50 = 2.4\degr\), which defines a thin radial wedge from
  \(C_1\).  The results are shown as red and blue dots superimposed
  on the images in Figure~\ref{fig:mipsgal-examples}. Each of the two
  methods, ``peak'' and ``mean'', works better in some objects and
  worse in others (according to the subjective judgment of
  ``correctly'' tracing the bow shock shape).  We therefore take the
  average by amalgamating all the \(C_{k,\text{peak}}\) and
  \(C_{k,\text{mean}}\) points into a single set, \(C_{k}\), for
  the following steps.
\item For each of the points \(C_{k}\), determine the radial
  separation from the central source, \(R_k = |C_k - C_0|\) and the
  angle from the bow shock axis about the central source
  \(\theta_k = \text{PA}(C_k, C_0) - \text{PA}_0\).  These are plotted in
  the upper left panels of Figure~\ref{fig:mipsgal-examples}.  Note
  that, even though the rays are cast from the step-back point \(C_1\)
  within \(\pm 60\degr\) of \(\text{PA}_0\), the angles \(\theta_k\) are
  measured from the source, \(C_0\), which is closer to the bow shock
  than \(C_1\) and therefore \(|\theta_k|\) can be much larger than
  \(60\degr\).
\item Make our own estimate of the axial size, \(R_0\), of the bow
  shock by calculating the mean and standard deviation of \(R_k\) over
  all points \(C_k\) with \(|\theta_k| \le 10\degr\).  Note that this is
  distinct from the nominal value of \(R_0\) given in the
  \citet{Kobulnicky:2016a} catalog, which was ``measured by eye''.
\item Estimate the radius of curvature, \(R_c\), by fitting a circle
  to all those points within \(\pm 45\degr\) of the nominal axis
  (\(|\theta_k| < 45\degr\)), but after excluding any point with
  \(R_k < \frac12 R_m\) or \(R_k > 2 R_m\), where \(R_m\) is the median
  \(R_k\) for \(|\theta_k| < 45\degr\).
\item Determine two separate estimates, \(R_{90+}\) and \(R_{90-}\),
  of the perpendicular radius, \(R_{90}\), by taking the mean and
  standard deviation of \(R_k\) over all points \(C_k\) with
  \(|\theta_k - 90\degr| \le 10\degr\) for \(R_{90+}\), and with
  \(|\theta_k + 90\degr| \le 10\degr\) for \(R_{90-}\).
\end{enumerate}

@@

\begin{figure*}
  \centering
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.45\textwidth, trim=0 0 30 0]
    {figs/mipsgal-Rc-R90-zoom-annotated}
        & \includegraphics[width=0.45\textwidth, trim=0 0 30 0]
          {figs/mipsgal-Rc-R90-thumbnails} 
  \end{tabular}
  \caption[]{MIPSGAL sources on the bow shock shape diagnostic diagram
    of dimensionless radius of curvature versus perpendicular radius.
    The regions corresponding to different classes of cuadrics are
    shown by shading (see \S~\ref{sec:conic}): oblate spheroids (light
    gray background); prolate spheroids (darker gray background);
    paraboloids (curved black line); hyperboloids (white background).
    (a) Individual sources with bow shock fit quality rating of 3-star
    or above.  All 5-star sources plus those 4-star sources with
    \(R_c/R_0 < 1\) are labelled with their \cite{Kobulnicky:2016a}
    catalog number.  Horizontal error bars do not directly reflect the
    uncertainty in \(R_c/R_0\) but are instead simply the standard
    deviation from the circle fit of bowshock points \(R_k\) within
    \(\pm 45\degr\) of the axis.  Values on the vertical axis
    represent the average of \(R_{90+}\) and \(R_{90-}\), with thin
    vertical error bars showing the difference between \(R_{90+}\) and
    \(R_{90-}\), and thick vertical error bars showing the rms
    dispersion of \(R_k\) about these values for bow shock points
    within \(\pm 10\degr\) of the \(+90\degr\) and \(-90\degr\)
    directions.  (b) Kernel density estimator (KDE) of the
    distribution for 3-star sources (blue, filled contours) and 4-
    plus 5-star sources (orange/brown, unfilled contours).  The KDE
    uses an anisotropic gaussian kernel with bandwidths of
    \(0.18 \times 0.12\). Images of representative 4- and 5-star sources at
    different points on the \(R_c\)--\(R_{90}\) plane are also shown.}
  \label{fig:mipsgal-shapes}
\end{figure*}

@@

\begin{figure*}
  \centering
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-environment} &
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-candidates} 
  \end{tabular}
  \caption[]{Example comparisons between the distribution of bowshock
    shapes when the sources are divided into two or more sub-samples
    according to the value of a categorical parameter.  Contours show
    the KDE of the distribution of each sub-sample for all 3-, 4-, and
    5-star sources, while filled circles show 5-star sources
    only. Inset histograms show the marginal distributions on the two
    shape axes. (a)~Source environment, divided into three
    sub-samples: ``Isolated'' (purple symbols and purple filled
    contours), ``Facing \hii{} region or \SI{8}{\um} bright-rimmed
    cloud'' (orange symbols and orange-brown hollow continuous
    contours), and ``Within \hii{} region'' (green symbols and green
    hollow dashed contours). (b)~Uncertainty in stellar source
    identification, divided into two sub-samples: ``Multiple
    candidates for stellar source'' (green symbols and green filled
    contours) and ``Single candidate for stellar source'' (orange
    symbols and orange-brown hollow continuous contours).}
  \label{fig:mipsgal-uncorrelated}
\end{figure*}

@@

\begin{figure}
  \centering
  \setlength\tabcolsep{0pt}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[height=1.167\linewidth]
    {figs/mipsgal-boxplot-Rc-R90-versus-R0}
        & \includegraphics[height=1.167\linewidth]
          {figs/mipsgal-boxplot-Rc-R90-versus-H0}
  \end{tabular}
  \caption{(a) Box plots of the distributions of bow shock shape
    parameters, \(R_c/R_0\) and \(R_{90} / R_0\), together with a
    measure of the tail asymmetry, \(|R_{90+} - R_{90-}| / R_{90}\),
    after partitioning on the bow shock size, \(R_0\).  All 3-, 4-,
    and 5-star sources are sorted according to \(R_0\) and divided
    into 5 non-overlapping sub-samples of roughly equal size, each
    labelled by their median value of \(R_0\) in arcseconds, as
    illustrated in the lower panel.  The colored boxes show the
    interquartile range, with the median indicated by a horizontal
    line, the 1st-to-9th interdecile range by error bars, and outliers
    by diamonds.  A 5-sample Anderson--Darling test is performed for
    each dependent variable, with resultant \(p\)-value given at the
    bottom of each panel.  It is apparent that the dispersion in
    \(R_{90} / R_0\) (upper panel) increases systematically with
    \(R_0\), although the central value is roughly constant.  No clear
    systematic changes are apparent in \(R_c / R_{90}\) (middle
    panel).  (b)~As (a), but partitioning on the extinction-corrected
    \(H\)-band magnitude, \(H_0\).  Note that sub-samples are plotted
    in order of \textit{decreasing} magnitude to allow comparison
    between (a) and (b), given the negative correlation between
    \(H_0\) and \(R_0\) (see Fig.~\ref{fig:mipsgal-pairplot}).}
  \label{fig:mipsgal-boxplot}
\end{figure}

@@

We also tested all the other source parameters listed in
\citeauthor{Kobulnicky:2016a}'s catalog, finding no significant shape
differences for sources with/without \SI{8}{\um} emission, with
low/high extinction, closer/farther from the Galactic plane, or
closer/farther from the Galactic center. Details are given in
Table~\ref{tab:big-p}.  In all these cases, differences in mean or
median \(R_{90}/R_0\) less than \(0.06\) and in \(R_c/R_0\) less than
\(0.16\) are found, which are not significant even at the 2-\(\sigma\)
level.

The only parameters that \emph{do} show a possible correlation with
the bowshock shape are the bow shock angular size, \(R_0\), and the
extinction-corrected magnitude of the central star, \(H_0\), which are
illustrated in Figure~\ref{fig:mipsgal-correlated}.

>>>>

We developed a methodology for automatically tracing the arcs as follows:
\begin{enumerate}[1.]
\item Calculate arrays of celestial coordinates, \(C\), for each pixel
  of the image.
\item Using the central source coordinates, \(C_0\) and nominal
  bowshock radius, \(R_0\) from \citet{Kobulnicky:2016a}, construct a
  pixel mask that includes only those pixels with separations from the
  source that satisfy \(\frac12 R_0 \le |C - C_0| \le 3 R_0\).  This mask
  will be used for all subsequent operations, which serves to help
  avoid confusion from the star itself and other bright sources in the
  field of view.
\item Define a ``step-back'' point, \(C_1\), which is located at a
  separation \(2 R_0\) from the source, but in the opposite direction
  from the apex of the bow shock. That is, along a position angle
  180\degr{} from the nominal position angle, \(\text{PA}_0\), of the
  bow shock axis.  This point is at one end of the orange line shown
  superimposed on the bow shock images in
  Figure~\ref{fig:mipsgal-examples}.
\item Looping over a grid of 50 position angles, \(\text{PA}_k\),
  within \(\pm 60\degr\) of \(\text{PA}_0\), estimate the location of
  the arc along rays cast from the step-back point, using two
  different methods:
  \begin{enumerate}[(a)]
  \item The pixel with the peak brightness, with coordinates
    \(C_{k,\text{peak}}\) (red dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \item The mean brightness-weighted separation from \(C_1\), with
    coordinates \(C_{k,\text{mean}}\) (light blue dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \end{enumerate}
  For each \(\text{PA}_k\) in the grid, the calculation is performed
  over only those pixels that satisfy
  \(|\text{PA}(C, C_1) - \text{PA}_k| < \frac12 \delta\theta\), where
  \(\delta\theta = 120/50 = 2.4\degr\), which defines a thin radial wedge from
  \(C_1\).  The results are shown as red and blue dots superimposed
  on the images in Figure~\ref{fig:mipsgal-examples}. Each of the two
  methods, ``peak'' and ``mean'', works better in some objects and
  worse in others (according to the subjective judgment of
  ``correctly'' tracing the bow shock shape).  We therefore take the
  average by amalgamating all the \(C_{k,\text{peak}}\) and
  \(C_{k,\text{mean}}\) points into a single set, \(C_{k}\), for
  the following steps.
\item For each of the points \(C_{k}\), determine the radial
  separation from the central source, \(R_k = |C_k - C_0|\) and the
  angle from the bow shock axis about the central source
  \(\theta_k = \text{PA}(C_k, C_0) - \text{PA}_0\).  These are plotted in
  the upper left panels of Figure~\ref{fig:mipsgal-examples}.  Note
  that, even though the rays are cast from the step-back point \(C_1\)
  within \(\pm 60\degr\) of \(\text{PA}_0\), the angles \(\theta_k\) are
  measured from the source, \(C_0\), which is closer to the bow shock
  than \(C_1\) and therefore \(|\theta_k|\) can be much larger than
  \(60\degr\).
\item Make our own estimate of the axial size, \(R_0\), of the bow
  shock by calculating the mean of \(R_k\) over all points \(C_k\)
  with \(|\theta_k| \le 10\degr\).  Note that this is distinct from the
  nominal value of \(R_0\) given in the \citet{Kobulnicky:2016a}
  catalog, which was ``measured by eye''.  We denote by
  \(\epsilon(R_0)\) the standard deviation of the \(R_k\) that go into
  calculating \(R_0\).\label{step:R0}
\item Estimate the radius of curvature, \(R_c\), by fitting a circle
  to all those points within \(\pm 45\degr\) of the nominal axis
  (\(|\theta_k| < 45\degr\)), but after excluding any point with
  \(R_k < \frac12 R_m\) or \(R_k > 2 R_m\), where \(R_m\) is the median
  \(R_k\) for \(|\theta_k| < 45\degr\).
\item Determine two separate estimates, \(R_{90+}\) and \(R_{90-}\),
  of the perpendicular radius, \(R_{90}\), by taking the mean of
  \(R_k\) over all points \(C_k\) with
  \(|\theta_k - 90\degr| \le 10\degr\) for \(R_{90+}\), and with
  \(|\theta_k + 90\degr| \le 10\degr\) for \(R_{90-}\).  The average of the
  two standard deviations of the \(R_k\) that contribute to
  \(R_{90+}\) and \(R_{90-}\) is denoted by \(\epsilon(R_{90})\).\label{step:R90}
\end{enumerate}

@@

\begin{figure*}
  \centering
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.45\textwidth, trim=0 0 30 0]
    {figs/mipsgal-Rc-R90-zoom-annotated}
        & \includegraphics[width=0.45\textwidth, trim=0 0 30 0]
          {figs/mipsgal-Rc-R90-thumbnails} 
  \end{tabular}
  \caption[]{MIPSGAL sources on the bow shock shape diagnostic diagram
    of dimensionless radius of curvature versus perpendicular radius.
    The regions corresponding to different classes of cuadrics are
    shown by shading (see \S~\ref{sec:conic}): oblate spheroids (light
    gray background); prolate spheroids (darker gray background);
    paraboloids (curved black line); hyperboloids (white background).
    (a) Individual sources with bow shock fit quality rating of 3-star
    or above.  All 5-star sources plus those 4-star sources with
    \(R_c/R_0 < 1\) are labelled with their \cite{Kobulnicky:2016a}
    catalog number.  Horizontal error bars do not directly reflect the
    uncertainty in \(R_c/R_0\) but are instead simply the standard
    deviation from the circle fit of bowshock points \(R_k\) within
    \(\pm 45\degr\) of the axis.  Values on the vertical axis
    represent the average of \(R_{90+}\) and \(R_{90-}\), with thin
    vertical error bars showing the difference between \(R_{90+}\) and
    \(R_{90-}\), and thick vertical error bars showing the rms
    dispersion of \(R_k\) about these values for bow shock points
    within \(\pm 10\degr\) of the \(+90\degr\) and \(-90\degr\)
    directions.  (b) Kernel density estimator (KDE) of the
    distribution for 3-star sources (blue, filled contours) and 4-
    plus 5-star sources (orange/brown, unfilled contours).  The KDE
    uses an anisotropic gaussian kernel with bandwidths of
    \(0.18 \times 0.12\). Thumbnail images of representative 4- and 5-star
    sources at different points on the \(R_c\)--\(R_{90}\) plane are
    also shown. The angular scale of each image is indicated by a
    yellow circle of diameter \(7.5''\), centered on the stellar
    source.}
  \label{fig:mipsgal-shapes}
\end{figure*}

@@

\begin{figure*}
  \centering
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-environment} &
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-candidates} 
  \end{tabular}
  \caption[]{Example comparisons between the distribution of bowshock
    shapes when the sources are divided into two or more sub-samples
    according to the value of a categorical parameter.  Contours show
    the KDE of the distribution of each sub-sample for all 3-, 4-, and
    5-star sources, while filled circle symbols show 5-star sources
    only. Inset histograms show the marginal distributions on the two
    shape axes. (a)~Source environment, divided into three
    sub-samples: ``Isolated'' (purple symbols and purple filled
    contours), ``Facing \hii{} region or \SI{8}{\um} bright-rimmed
    cloud'' (orange symbols and orange-brown hollow continuous
    contours), and ``Within \hii{} region'' (green symbols and green
    hollow dashed contours). (b)~Uncertainty in stellar source
    identification, divided into two sub-samples: ``Multiple
    candidates for stellar source'' (green symbols and green filled
    contours) and ``Single candidate for stellar source'' (orange
    symbols and orange-brown hollow continuous contours).}
  \label{fig:mipsgal-uncorrelated}
\end{figure*}

@@

\begin{figure}
  \centering
  \setlength\tabcolsep{0pt}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[height=1.167\linewidth]
    {figs/mipsgal-boxplot-Rc-R90-versus-R0}
        & \includegraphics[height=1.167\linewidth]
          {figs/mipsgal-boxplot-Rc-R90-versus-H0}
  \end{tabular}
  \caption{(a) Box plots of the distributions of bow shock shape
    parameters, \(R_c/R_0\) and \(R_{90} / R_0\), together with a
    measure of the tail asymmetry, \(|R_{90+} - R_{90-}| / R_{90}\),
    after partitioning on the bow shock size, \(R_0\).  All 3-, 4-,
    and 5-star sources are sorted according to \(R_0\) and divided
    into 5 non-overlapping sub-samples of roughly equal size, each
    labelled by their median value of \(R_0\) in arcseconds, as
    illustrated in the lower panel.  The colored boxes show the
    interquartile range, with the median indicated by a horizontal
    line, the 1st-to-9th interdecile range by error bars, and outliers
    by diamonds.  A 5-sample Anderson--Darling test and Brown--Forsythe
    test is performed for each dependent variable, with resultant
    \(p\)-value given at the bottom of each panel.   It is apparent
    that the dispersion in \(R_{90} / R_0\) (upper panel) increases
    systematically with \(R_0\), although the central value is roughly
    constant.  No clear systematic changes are apparent in
    \(R_c / R_{90}\) (middle panel).  (b)~As (a), but partitioning on
    the extinction-corrected \(H\)-band magnitude, \(H_0\).  Note that
    sub-samples are plotted in order of \textit{decreasing} magnitude
    to allow comparison between (a) and (b), given the negative
    correlation between \(H_0\) and \(R_0\) (see
    Fig.~\ref{fig:mipsgal-pairplot}).}
  \label{fig:mipsgal-boxplot}
\end{figure}

@@

We also tested all the other source parameters listed in
\citeauthor{Kobulnicky:2016a}'s catalog, finding no significant shape
differences for sources with/without \SI{8}{\um} emission, with
low/high extinction, closer/farther from the Galactic plane, or
closer/farther from the Galactic center. Details are given in
Table~\ref{tab:big-p}.  In all these cases, differences in mean or
median \(R_{90}/R_0\) less than \(0.06\) and in \(R_c/R_0\) less than
\(0.16\) are found, corresponding to rank biserial correlation
coefficients (a non-parametric dimensionless measure of the difference
between two samples, see Appendix~\ref{sec:distr-p-values}) of
\(r_b < 0.15\), which are not significant even at the 2-\(\sigma\) level.

The only parameters that \emph{do} show a possible correlation with
the bowshock shape are the bow shock angular size, \(R_0\), and the
extinction-corrected magnitude of the central star, \(H_0\), which are
illustrated in Figure~\ref{fig:mipsgal-correlated}.  In both cases,
the general purpose Anderson--Darling test indicates differences in the
sub-sample distributions of \(R_{90} / R_0\) at the 2-\(\sigma\) level.  It
is apparent from the histograms shown in the right-hand inset graphs
of Figure~\ref{fig:mipsgal-correlated}a that in the case of the
small/large \(R_0\) sub-samples it is the width rather than the
central tendency of the distributions that is different, which is
confirmed by the more specific Brown--Forsythe test, which indicates a
difference between the sub-sample dispersions at the 3-\(\sigma\) level.
In the case of the faint/bright \(H_0\) sub-samples
(Fig.~\ref{fig:mipsgal-heteroscedastic}b), it is less clear what
feature of the distributions differ.

In order to investigate these effects in more detail and look for
systematic trends, we divide the independent parameter (\(R_0\) or
\(H_0\)) into 5 rather than 2 equal-sized sub-samples, with results
shown as box plots in Figure~\ref{fig:mipsgal-boxplot}.  A systematic
increase with \(R_0\) in the dispersion of \(R_{90}/R_0\) (upper panel
of Fig.~\ref{fig:mipsgal-boxplot}a) is apparent from both the
interquartile range (colored boxes) and interdecile range (error
bars).  As a check on whether observational uncertainties might be
contributing to this trend, the third and fourth rows of box plots
show the statistics for, respectively, the fractional asymmetry of the
bowshock wings and the standard deviation, \(\epsilon(R_{90})\), of the
individual points on the bow shock that go into determining \(R_{90}\)
(see step~\ref{step:R90} of the tracing/fitting methodology described
in \S~\ref{sec:autom-trac-fitt}).  It can be seen that neither of
these quantities tends to increase with \(R_0\), and in fact there is
a significant \emph{decrease} in \(\epsilon(R_{90})\) with \(R_0\).  This
implies that the increase with angular size of the diversity of bow
shock wing shapes is real, and not due to observational uncertainties.  
On the other hand, the fact that no clear trends are evident as a
function of source magnitude \(H_0\) (Fig.~\ref{fig:mipsgal-boxplot}b)
leads us to suspect that the \(p < 0.05\) result obtained for the
2-sample Anderson--Darling test was a \emph{false positive}, which is
also supported by the analysis in Appendix~\ref{sec:distr-p-values}
and Figure~\ref{fig:histo-p-values}.

So, the only correlation that survives scrutiny is the increase in
dispersion of \(R_{90}/R_0\) with increasing \(R_0\), which is finally
presented in the greatest detail in
Figure~\ref{fig:mipsgal-heteroscedastic}, where eschew any binning and
perform linear regressions of the individual sources as a function of
\(\log_{10} R_0\).  As expected from the previous 2-sample and
5-sample analyses, the regression of \(R_{90}/R_0\) is almost flat,
with a low correlation coefficient.
